% !TEX root = ../main/main.tex
% Acronyms and Abbreviations (front-matter insert)
\chapter*{Acronyms}
\addcontentsline{toc}{chapter}{Acronyms}
\begingroup\RaggedRight\sloppy
\begin{description}
  \item[ATS] Against the Spread.
  \item[CBV] Comparative Book Value. Difference between model-implied fair price and posted price, adjusted for hold.
  \item[CLV] Closing Line Value. Improvement (bps) of our executed price vs the market closing price.
  \item[COPULA] A function that couples univariate marginals into a multivariate distribution; used here for spread–total dependence.
  \item[CQL] Conservative Q-Learning (offline RL algorithm).
  \item[CRPS] Continuous Ranked Probability Score (proper scoring rule for distributions).
  \item[CVaR] Conditional Value at Risk (expected tail loss beyond VaR).
  \item[DR] Doubly Robust estimator (off-policy evaluation).
  \item[ECE] Expected Calibration Error.
  \item[ESS] Effective Sample Size.
  \item[HCOPE] High-Confidence Off-Policy Evaluation (lower bounds).
  \item[IQL] Implicit Q-Learning (offline RL algorithm).
  \item[KF] Kalman Filter (linear-Gaussian state estimation).
  \item[MAR] MAR ratio (mean/absolute drawdown or similar risk-adjusted return, as defined in Chapter~\ref{chap:results}).
  \item[OPE] Off-Policy Evaluation.
  \item[PIT] Probability Integral Transform.
  \item[PROE] Pass Rate Over Expected.
  \item[RF] Random Forest (used in live WP optional scaffolding).
  \item[ROI] Return on Investment.
  \item[SGP] Same-Game Parlay.
  \item[SNIS] Self-Normalized Importance Sampling (off-policy evaluation).
  \item[TD3+BC] TD3 with Behavior Cloning regularization (offline RL algorithm).
\end{description}
\endgroup
