---
title: "01 – Ingest Schedules"
format:
  html:
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
params:
  start_season: 1999
  end_season: 2024
  seasons_override: null
  snapshot_path: "data/raw/nflverse_schedules_1999_2024.rds"
  truncate_games: true
---

## Purpose
Document the end-to-end process that refreshes the `games` table with nflverse schedules and embedded odds fields (moneylines, spread odds, totals) for 1999-2024 seasons.

## Prerequisites
- `docker compose up -d pg`
- R libraries from `renv.lock` (`nflreadr`, `dplyr`, `DBI`, `RPostgres`, `lubridate`)
- Environment variables from `.env` (if overriding database credentials)

## 1. Load Schedules Snapshot
```{r}
library(nflreadr)
library(dplyr)

sched_years <- if (!is.null(params$seasons_override)) params$seasons_override else params$start_season:params$end_season

schedules <- load_schedules(sched_years)
count_schedules <- schedules |> count(season)
count_schedules
```

## 2. Persist Raw Snapshot
```{r}
readr::write_rds(schedules, params$snapshot_path)
```

## 3. Transform For Warehouse
```{r}
library(lubridate)

games_table <- schedules |>
  transmute(
    game_id,
    season,
    week,
    home_team,
    away_team,
    kickoff      = ymd(gameday),
    spread_close = spread_line,
    total_close  = total_line,
    home_score,
    away_score,
    home_moneyline,
    away_moneyline,
    home_spread_odds,
    away_spread_odds,
    over_odds,
    under_odds
  )

nrow(games_table)
```

## 4. Load Into Postgres
```{r}
library(DBI)
library(RPostgres)

con <- dbConnect(
  Postgres(),
  dbname   = Sys.getenv("POSTGRES_DB", "devdb01"),
  host     = Sys.getenv("POSTGRES_HOST", "localhost"),
  port     = as.integer(Sys.getenv("POSTGRES_PORT", 5544)),
  user     = Sys.getenv("POSTGRES_USER", "dro"),
  password = Sys.getenv("POSTGRES_PASSWORD", "sicillionbillions")
)

dbWithTransaction(con, {
  if (isTRUE(params$truncate_games)) {
    dbExecute(con, "TRUNCATE TABLE games;")
  }
  dbWriteTable(con, "games", games_table, append = TRUE, row.names = FALSE)
})

dbDisconnect(con)
```

## 5. Sanity Checks
```{r}
library(DBI)
con <- dbConnect(
  Postgres(),
  dbname   = Sys.getenv("POSTGRES_DB", "devdb01"),
  host     = Sys.getenv("POSTGRES_HOST", "localhost"),
  port     = as.integer(Sys.getenv("POSTGRES_PORT", 5544)),
  user     = Sys.getenv("POSTGRES_USER", "dro"),
  password = Sys.getenv("POSTGRES_PASSWORD", "sicillionbillions")
)

dbGetQuery(con, "SELECT season, COUNT(*) AS games FROM games GROUP BY season ORDER BY season LIMIT 10;")

dbDisconnect(con)
```

## TODO
- Extend loader to incorporate alternative data sources for pre-1999 seasons if secured.
