---
title: "02 – Play-by-Play Features"
format:
  html:
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
params:
  season_start: 2023
  season_end: 2024
  seasons_override: null
  write_raw_pbp: false
  write_team_mart: true
  mart_schema: "mart"
  truncate_team_mart: false
  persist_parquet: true
  parquet_path: "data/raw/team_epa_{season_start}_{season_end}.parquet"
---

## Purpose
Build play-level features and EPA aggregates to support modeling and matchup reports.

## Prerequisites
- Postgres container running with `plays` schema applied (`db/001_init.sql`).
- R packages: `nflreadr`, `dplyr`, `purrr`, `dbplyr`, `DBI`, `RPostgres`, `glue`, `arrow`.

## 1. Load Raw PBP Data
```{r}
library(nflreadr)
library(dplyr)
library(purrr)

pbp_years <- if (!is.null(params$seasons_override)) params$seasons_override else params$season_start:params$season_end

pbp_raw <- purrr::map_dfr(pbp_years, load_pbp)

pbp_raw |> select(game_id, play_id, posteam, defteam, epa) |> head()
```

## 2. Derive Aggregations
```{r}
team_epa <- pbp_raw |>
  filter(!is.na(epa)) |>
  group_by(game_id, posteam) |>
  summarise(
    plays = n(),
    epa_sum = sum(epa),
    epa_mean = mean(epa),
    explosive_pass = mean(air_yards >= 15, na.rm = TRUE),
    explosive_rush = mean(rush == 1 & yards_gained >= 10, na.rm = TRUE),
    .groups = "drop"
  )

tail(team_epa)
```

## 3. Write Aggregates to Warehouse
```{r}
library(DBI)
library(RPostgres)

con <- dbConnect(
  Postgres(),
  dbname   = Sys.getenv("POSTGRES_DB", "devdb01"),
  host     = Sys.getenv("POSTGRES_HOST", "localhost"),
  port     = as.integer(Sys.getenv("POSTGRES_PORT", 5544)),
  user     = Sys.getenv("POSTGRES_USER", "dro"),
  password = Sys.getenv("POSTGRES_PASSWORD", "sicillionbillions")
)

if (isTRUE(params$write_raw_pbp)) {
  dbWriteTable(
    con,
    "plays",
    pbp_raw |> select(game_id, play_id, posteam, defteam, quarter, game_seconds_remaining, down, ydstogo, epa, pass, rush),
    append = TRUE,
    row.names = FALSE
  )
}

if (isTRUE(params$write_team_mart)) {
  schema_name <- params$mart_schema
  quoted_schema <- as.character(DBI::dbQuoteIdentifier(con, schema_name))
  dbExecute(con, sprintf("CREATE SCHEMA IF NOT EXISTS %s;", quoted_schema))
  if (isTRUE(params$truncate_team_mart)) {
    dbExecute(con, sprintf("TRUNCATE TABLE %s.team_epa;", quoted_schema))
  }
  dbWriteTable(con, Id(schema = schema_name, table = "team_epa"), team_epa, append = TRUE, row.names = FALSE)
}

dbDisconnect(con)
```

## 4. Persist Feature Table Locally
```{r}
if (isTRUE(params$persist_parquet)) {
  output_path <- glue::glue(params$parquet_path, season_start = params$season_start, season_end = params$season_end)
  arrow::write_parquet(team_epa, output_path)
}
```

## TODO
- Add incremental load logic or upserts for raw play storage and mart tables.
- Join to `games` and roster data for opponent context prior to modeling.
- Emit basic QA metrics (coverage, mean EPA stability) after each refresh.
