---
title: "10 – Spread Model (XGBoost)"
format:
  html:
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
params:
  train_seasons: [2014, 2015, 2016, 2017, 2018, 2019, 2020]
  test_seasons: [2021, 2022, 2023]
  mart_schema: "mart"
  team_epa_table: "team_epa"
  outcome: "spread_target"
  trees: 500
  grid_size: 20
  export_metrics: false
  metrics_path: "models/spread_xgb_metrics.csv"
---

## Goal
Train and evaluate a gradient boosted model that predicts against-the-spread outcomes using enriched schedule, odds, and play-derived features. Seasons and model hyperparameters are configurable via parameters.

## Prerequisites
- Materialised EPA features stored in `{params$mart_schema}.{params$team_epa_table}`.
- R packages: `tidymodels`, `DBI`, `RPostgres`, `dplyr`, `glue`, `skimr`, `arrow` (optional for exports).

## 1. Assemble Training Data
```{r}
library(DBI)
library(RPostgres)
library(dplyr)
library(glue)

con <- dbConnect(
  Postgres(),
  dbname   = Sys.getenv("POSTGRES_DB", "devdb01"),
  host     = Sys.getenv("POSTGRES_HOST", "localhost"),
  port     = as.integer(Sys.getenv("POSTGRES_PORT", 5544)),
  user     = Sys.getenv("POSTGRES_USER", "dro"),
  password = Sys.getenv("POSTGRES_PASSWORD", "sicillionbillions")
)

team_table <- Id(schema = params$mart_schema, table = params$team_epa_table)

sql <- glue(""
SELECT
  g.season,
  g.week,
  g.game_id,
  g.home_team,
  g.away_team,
  g.home_score,
  g.away_score,
  g.spread_close,
  g.home_moneyline,
  g.away_moneyline,
  hepa.epa_mean AS home_epa_mean,
  aepa.epa_mean AS away_epa_mean
FROM games g
LEFT JOIN {DBI::dbQuoteIdentifier(con, params$mart_schema)}.{DBI::dbQuoteIdentifier(con, params$team_epa_table)} hepa
  ON g.game_id = hepa.game_id AND g.home_team = hepa.posteam
LEFT JOIN {DBI::dbQuoteIdentifier(con, params$mart_schema)}.{DBI::dbQuoteIdentifier(con, params$team_epa_table)} aepa
  ON g.game_id = aepa.game_id AND g.away_team = aepa.posteam
WHERE g.season BETWEEN {min(c(params$train_seasons, params$test_seasons))} AND {max(c(params$train_seasons, params$test_seasons))}
""")

df <- dbGetQuery(con, sql)

dbDisconnect(con)
```

## 2. Feature Engineering
```{r}
library(tidyr)
library(skimr)

df <- df |>
  mutate(
    margin = home_score - away_score,
    spread_target = if_else(margin + spread_close > 0, 1, 0)
  ) |>
  drop_na(spread_target)

skim(df)
```

## 3. Train/Test Split
```{r}
library(tidymodels)

train_df <- df |> filter(season %in% params$train_seasons)
test_df  <- df |> filter(season %in% params$test_seasons)

rec <- recipe(spread_target ~ spread_close + home_epa_mean + away_epa_mean + home_moneyline + away_moneyline, data = train_df) |>
  step_impute_knn(all_predictors()) |>
  step_normalize(all_predictors())

xgb_spec <- boost_tree(
  trees = params$trees,
  learn_rate = tune(),
  tree_depth = tune(),
  min_n = tune()
) |>
  set_engine("xgboost") |>
  set_mode("classification")

wf <- workflow() |> add_recipe(rec) |> add_model(xgb_spec)
```

## 4. Hyperparameter Tuning
```{r}
set.seed(42)
res <- wf |> tune_grid(
  resamples = vfold_cv(train_df, v = 5, strata = spread_target),
  grid = params$grid_size,
  metrics = metric_set(roc_auc, accuracy)
)

show_best(res, metric = "roc_auc")
```

## 5. Final Fit and Evaluation
```{r}
best_params <- select_best(res, metric = "roc_auc")
final_wf <- finalize_workflow(wf, best_params)

final_fit <- final_wf |> fit(train_df)

test_preds <- predict(final_fit, test_df, type = "prob") |>
  bind_cols(test_df)

metrics <- yardstick::metric_set(roc_auc, accuracy)(test_preds, truth = spread_target, .pred_1)
metrics

if (isTRUE(params$export_metrics)) {
  fs::dir_create(dirname(params$metrics_path))
  readr::write_csv(metrics, params$metrics_path)
}
```

## TODO
- Log feature importances and calibration diagnostics per model run.
- Compare against baseline heuristics (e.g., spread-only logistic regression).
- Persist fitted model objects for downstream deployment.
