---
title: "11 – Monte Carlo Skellam"
format:
  html:
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
params:
  season_filter: 2024
  sims: 50000
  seed: 42
  export_summary: false
  summary_path: "models/skellam_summary.parquet"
---

## Purpose
Simulate market-implied score distributions using Skellam assumptions and compare against closing lines/odds. Parameters allow season and simulation controls.

## Prerequisites
- Postgres `games` data with spreads and totals populated.
- R packages: `dplyr`, `DBI`, `RPostgres`, `purrr`, `ggplot2`, `arrow`, `glue`.

## 1. Retrieve Market Lines
```{r}
library(DBI)
library(RPostgres)
library(dplyr)
library(glue)

con <- dbConnect(
  Postgres(),
  dbname   = Sys.getenv("POSTGRES_DB", "devdb01"),
  host     = Sys.getenv("POSTGRES_HOST", "localhost"),
  port     = as.integer(Sys.getenv("POSTGRES_PORT", 5544)),
  user     = Sys.getenv("POSTGRES_USER", "dro"),
  password = Sys.getenv("POSTGRES_PASSWORD", "sicillionbillions")
)

season_clause <- if (is.null(params$season_filter)) "" else glue("WHERE season = {as.integer(params$season_filter)}")
lines <- dbGetQuery(con, glue("SELECT game_id, season, week, home_team, away_team, spread_close, total_close FROM games {season_clause}"))

dbDisconnect(con)
```

## 2. Derive Team Means From Spread/Total
```{r}
lines <- lines |>
  mutate(
    lambda_home = (total_close / 2) - (spread_close / 2),
    lambda_away = (total_close / 2) + (spread_close / 2)
  )
```

## 3. Monte Carlo Simulation
```{r}
library(purrr)
library(tidyr)
library(dplyr)
set.seed(params$seed)

simulate_game <- function(game_row, sims) {
  home_scores <- rpois(sims, game_row$lambda_home)
  away_scores <- rpois(sims, game_row$lambda_away)
  tibble::tibble(
    game_id = game_row$game_id,
    home_scores = home_scores,
    away_scores = away_scores,
    margin = home_scores - away_scores
  )
}

sim_results <- map_dfr(split(lines, lines$game_id), simulate_game, sims = params$sims)
```

## 4. Probability Estimates
```{r}
probabilities <- sim_results |>
  left_join(lines |> select(game_id, spread_close, total_close), by = "game_id") |>
  group_by(game_id, spread_close, total_close) |>
  summarise(
    cover_prob = mean(margin > -spread_close),
    push_prob  = mean(margin == -spread_close),
    total_over = mean(home_scores + away_scores > total_close),
    .groups = "drop"
  )

probabilities
```

## 5. Visualize Distributions
```{r}
if (nrow(sim_results) > 0) {
  example_game <- sim_results |> filter(game_id == sample(unique(game_id), 1)) |>
    left_join(lines |> select(game_id, spread_close), by = "game_id")

  ggplot(example_game, aes(x = margin)) +
    geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
    geom_vline(xintercept = -unique(example_game$spread_close), color = "red", linetype = "dashed") +
    labs(title = "Simulated Margin Distribution", x = "Home Margin", y = "Frequency")
}
```

## 6. Optional Export
```{r}
if (isTRUE(params$export_summary)) {
  arrow::write_parquet(probabilities, params$summary_path)
}
```

## TODO
- Incorporate team-specific scoring rates (EPA-derived lambdas) instead of pure market averages.
- Compare simulated probabilities vs. actual outcomes to quantify edge.
- Surface distributions via dashboards (Quarto/Plotly).
